Introducing statistics with \\texttt{intRo}

### 2014-12-05 student-paper-competition-2014/paper.Rmd
tweaked conclusion

@@

We have three fundamental philosophies that guided the creation of \texttt{intRo}, and this comes out in the design. Primarily, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics. Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for his or her own classroom needs.

@@

Within the data category there are two modules, sources and transformation. The data sources module allows the user to select or upload a dataset. Several datasets are pre-included, which can be selected via a drop-down menu. The user can check a box to upload a dataset, and then select the CSV file from his computer. Upon selection or successful upload of a dataset, a table view of the data will be visible on the right. From this module the user can select a subset of the data either by value or random selection. The dataset selected in the data module will be made available to all other modules of \texttt{intRo}. The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric. Additionally, numeric variable can be transformed using a power transformation. The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.

@@

Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in his first semester. The contingency table module allows the user to compute a contingency table for two categorical variables. The results can either be displayed as counts, row percentages, column percentages, or total percentages. The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x). The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression. Results of the regression will be displayed on the right. A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed. A user also has the ability to save residuals and predicted values to their dataset for further analysis. The T test module allows the user to perform either a one variable or two variable Student's T test. The user can select configuration options, including the direction of the test and the confidence level of the test. The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.

@@

Statistics courses with a lab component can most benefit from the use of \texttt{intRo}. As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries. They simply upload a dataset and click the module appropriate to the section of the class. Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in. Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results. For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available.

@@

We believe \texttt{intRo} can be a powerful and effective tool for introductory statistics education. Its modular structure allows it to be flexible enough for many different applications and curriculums. Its ease-of-use allows the student to focus his or her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation. Reproducible code generated from each analysis can be used to spark an interest in \texttt{R} programming in those who might otherwise not be exposed to it.

Despite the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more accessible and useful. In particular, we have begun development on an \texttt{R} package which will allow modules to be created automatically. This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo}. This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curriculums, including more advanced statistics courses. We will also be trialing \texttt{intRo}'s use in a statistics course at Iowa State University in Spring Semester 2015. This will allow us to assess its usefulness relative to software used in the past, as well as areas for improvement. Challenges exist with regards to the wider adoption of \texttt{intRo}. For instance, we will need to monitor how well the server \texttt{intRo} is hosted on handles the load of dozens of students performing data analyses. If performance issues are noted, the infrastructure used may need to be expanded to handle current and future load.

>>>>

\ak{[Add some references about software discouraging people, and mention other new tools out there like swirl.]}

@@

We have three fundamental philosophies that guided the creation of \texttt{intRo}, and this comes out in the design. Primarily, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics. Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for her own classroom needs.

@@

Within the data category there are two modules, sources and transformation. The data sources module allows the user to select or upload a dataset. Several datasets are pre-included, which can be selected via a drop-down menu. The user can check a box to upload a dataset, and then select the CSV file from her computer. Upon selection or successful upload of a dataset, a table view of the data will be visible on the right. From this module the user can select a subset of the data either by value or random selection. The dataset selected in the data module will be made available to all other modules of \texttt{intRo}. The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric. Additionally, numeric variable can be transformed using a power transformation. The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.

@@

Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester. The contingency table module allows the user to compute a contingency table for two categorical variables. The results can either be displayed as counts, row percentages, column percentages, or total percentages. The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x). The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression. Results of the regression will be displayed on the right. A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed. A user also has the ability to save residuals and predicted values to their dataset for further analysis. The T test module allows the user to perform either a one variable or two variable Student's T test. The user can select configuration options, including the direction of the test and the confidence level of the test. The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.

@@

Statistics courses with a lab component can most benefit from the use of \texttt{intRo}. As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries. They simply upload a dataset and click the module appropriate to the section of the class. Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in. Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results. For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within \texttt{intRo} as detailed in section \ref{sec:extensible}.

@@

We believe \texttt{intRo} can be a powerful and effective tool for introductory statistics education. Its modular structure allows it to be flexible enough for many different applications and curriculums. Its ease-of-use allows the student to focus her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation. Reproducible code generated from each analysis can be used to spark an interest in \texttt{R} programming in those who might otherwise not be exposed to it.

In addition to the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more useful to both students and instructors. In particular, we have begun development on an \texttt{R} package which will allow \texttt{intRo} modules to be created automatically from user written \texttt{R} code. This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo} as well as translate user code to \texttt{intRo} compatible code as populates the necessary files. This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.

Additionally,  \texttt{intRo} will be trialed in an introductory statistics course at Iowa State University in Spring 2015. This will allow us to assess its usefulness relative to software used in the past, as well as guage areas for improvement. Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in \texttt{R}.

Challenges do exist with regards to the wider adoption of \texttt{intRo}. For instance, we will need to monitor how well the server hosting \texttt{intRo} handles the load of dozens of students performing data analyses at once. If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load. An unknown quantity will be how feasible it is to increase adoption of \texttt{intRo} across the department, as well as to other universities.

Regardless, tools that focus on usability and extensibility in statistics education, such as \texttt{intRo}, are sure to encourage the next round of innovators to be interested and excited about statistical computing.
